---
title: "lab8_DerekSitu"
author: "Derek Situ (62222807)"
date: "3/13/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data
```{r}
dat <- read.delim("souvenir.txt", header = FALSE)
sales <- ts(dat$V1, start = c(1987, 1), frequency = 12)
```

## Question 1
```{r}
plot(sales)
acf(sales)
```
From the plot of the series, we notice a seasonal pattern, and we notice that the variation increases over time. This fits a multiplicative model. The acf also displays the seasonal pattern. There also may be a trend.

## Question 2
```{r}
training <- window(sales, start = c(1987, 1), end = c(1992, 12))
hw_fit <- HoltWinters(training, seasonal = "multiplicative")

c(hw_fit$alpha, hw_fit$beta, hw_fit$gamma)
```
The parameters are $\alpha = 0.347, \beta = 0.075, \gamma = 0.571, p = 12$.

``` {r}
plot(hw_fit)
```

## Question 3
```{r}
predict_test <- predict(hw_fit, 12, prediction.interval = TRUE, level = 0.95)
plot(predict_test[1:12], type = "l")
lines(predict_test[13:24], type = "c", col = 4) # plot upper bound
lines(predict_test[25:36], type = "c", col = 4) # plot lower bound
legend("topleft", lty = 1, col = c(1, 4),
       legend = c("Predicted values", "95% confidence interval"))

predict_test[1:3] # predicted values for first 3 months
```
The forecast values for the first 3 months of 1993 are `r predict_test[1:3]`.

## Question 4
```{r}
data.frame(Observed = sales[73:75], CI_Upper = predict_test[13:15],
           CI_Lower = predict_test[25:27]) |>
             dplyr::mutate(Observed_in_CI = 
                      ((Observed > CI_Lower) & (Observed < CI_Upper)))
```
Only the observed value for the first month of 1993 fits inside the 95% confidence interval.

## Question 5
I would consider a log transformation since the series seems to be increasing exponentially. One useful property of logs is that the log of a product is the sum of the logs of the factors, so using this property we can change this multiplicative model into an additive model.

