---
title: "Lab 2"
author: "Derek Situ (62222807)"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/dsder/OneDrive/Desktop/STAT_443/Lab_2")
library(zoo)
```

### Question 1
## (1a)
```{r, echo = TRUE}
# read in and plot the data
temperature_data <- read.csv("dataTempPG.csv", header = TRUE)
mean_min <-
  ts(temperature_data$Annual,
     start = c(1919, 1),
     frequency = 1)
plot(mean_min,
     main = "Prince George annual mean minimum temperatures",
     xlab = "Year",
     ylab = "Degrees (C)")
```
This series trends upwards. While there is evident local variation, a constant
seasonal period cannot be easily identified.

## (1b)
```{r, echo = TRUE}
# plot acf function
acf(mean_min, lag.max = 89)
```
We see that the acf slowly decays to 0 as the lag increases because of the 
trend. At lag = 36, the acf flips over the x-axis to the negative range.

## (1c)
```{r, echo = TRUE}
# plot 5-year rolling mean on top of 1968 to 2008 window
mean_min_1968_2008 <- window(mean_min, start = 1968, end = 2008)
plot(mean_min_1968_2008,
     main = "Prince George annual mean minimum temperatures",
     xlab = "Year",
     ylab = "Degrees (C)")
lines(rollmean(mean_min_1968_2008, k = 5), col = 2)
legend(
  "topleft",
  legend = c("Annual temperature", "5-year rolling mean"),
  lty = 1,
  col = c(1, 2)
)
```

### Question 2
## (2a)
```{r, echo = TRUE}
# read in data and plot acf
lake_data <- read.csv("LakeLevels.csv", header = TRUE)
depth <-
  ts(lake_data$LakeLevel,
     start = c(2007, 1),
     frequency = 365)
acf(depth, lag.max = 1800)
```
The acf oscillates around 0 since the series oscillates, and approaches 0 as the 
lag increases.

## (2b)
```{r, echo = TRUE}
# plot moving average decomposition
depth_decomp <- decompose(depth, type = "additive")
plot(depth_decomp)
```

## (2c)
```{r, echo = TRUE}
# plot loess decomposition
depth_loess <- stl(depth, s.window = "periodic")
plot(depth_loess)
```
 